---
title: "30-Day Mortality Rate of Myocardia Infarction Patients Admitted to All Seven Care Units"
author: Yu Chen 305266880
output:
  # ioslides_presentation: default
  html_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

# Data Preparation

Load database libraries and the tidyverse frontend:
```{r}
library(DBI)
library(RPostgreSQL)
library(tidyverse)
library(lubridate)
```

Credentials for using PostgreSQL database. We are going to use username `postgres` with password `postgres` to access the `mimic` database in the schemee `mimiciii`. 
```{r}
# Load configuration settings
dbdriver <- 'PostgreSQL'
#host  <- '127.0.0.1'
#port  <- '5432'
user  <- 'postgres'
password <- 'postgres'
dbname <- 'mimic'
schema <- 'mimiciii'
# Connect to the database using the configuration settings
con <- dbConnect(RPostgreSQL::PostgreSQL(), 
                 dbname = dbname, 
                 #host = host, 
                 #port = port, 
                 user = user, 
                 password = password)
# Set the default schema
dbExecute(con, paste("SET search_path TO ", schema, sep=" "))
con
```

In this section, we demo how to create a cohort of patients admitted into CCU, CSRU, MICU, NICU, NWARD, SICU and TSICU and were diagnosed with Myocardia Infarction.

First we create a (query) table of patients who were directly admitted into CCU, CSRU, MICU, NICU, NWARD, SICU and TSICU.
```{r}
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "CCU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> ccu_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "CSRU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> csru_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "MICU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> micu_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "NICU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> nicu_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "NWARD") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> nward_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "SICU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> sicu_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "TSICU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> tsicu_admissions
```

Now we want to restrict to Myocardia Infarction Patients patients. To find all possible ICD-9 codes related to heart attack, we search for string `myocardial infarction` in the `long_title` of table `d_icd_diagnoses`:
```{r}
mi_codes <- tbl(con, "d_icd_diagnoses") %>%
  filter(str_detect(tolower(long_title), "myocardial infarction")) 
```

`diagnoses_icd` table stores the diagnosis of each admission. We use `semi_join()` to keep the rows in `diagnoses_icd` that match the ICD-9 codes related to fever:
```{r}
tbl(con, "diagnoses_icd") %>%
  semi_join(mi_codes, by = "icd9_code") %>%
  print() -> mi_admissions
```

MI may not be listed as the principal diagnosis; as explained in [the documentation for the `patients` table](https://mimic.physionet.org/mimictables/diagnoses_icd/), the `seq_num` field is a priority ranking for the diagnoses generated at the end of stay. In order to focus on patients for whom MI was central to their hospitalization, we will include records with MI in any of the first five diagnosis positions, according to the `seq_num` field. To avoid duplicate admissions, we use `group_by()` and `top_n()` to limit the query to the first MI diagnosis for each admission.
```{r}
mi_admissions <- mi_admissions %>%
  filter(seq_num <= 5) %>%
  group_by(subject_id, hadm_id) %>%
  filter(min_rank(seq_num) <= 1) %>%
  ungroup() %>%
  select(subject_id, hadm_id, icd9_code, seq_num)
```

Now we `inner_join` the table of admissions to all the unit cares and the table of admissions that include MI diagnosis.
```{r}
study_admissions_ccu <- ccu_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_csru <- csru_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_micu <- micu_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_nicu <- nicu_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_nward <- nward_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_sicu <- sicu_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_tsicu <- tsicu_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
```

We notice that there is no data in table `study_admissions_nicu` and `study_admissions_nward`, which means there are no fever patients addmitted to NICU and NWARD. So we won't consider these two conditions in the following analysis.

Now we create a logical variable indicating the MI is the principal diagonosis or not (according to `seq_num`).
```{r}
study_admissions_ccu <- study_admissions_ccu %>%
  mutate(principal_dx = seq_num == 1) %>%
  select(-seq_num) %>%
  print()
study_admissions_csru <- study_admissions_csru %>%
  mutate(principal_dx = seq_num == 1) %>%
  select(-seq_num) %>%
  print()
study_admissions_micu <- study_admissions_micu %>%
  mutate(principal_dx = seq_num == 1) %>%
  select(-seq_num) %>%
  print()
study_admissions_sicu <- study_admissions_sicu %>%
  mutate(principal_dx = seq_num == 1) %>%
  select(-seq_num) %>%
  print()
study_admissions_tsicu <- study_admissions_tsicu %>%
  mutate(principal_dx = seq_num == 1) %>%
  select(-seq_num) %>%
  print()
```

We want to add information about the severity of patients’ ailments. The `drgcodes` table contains, for `DRG` codes from the All Payers Registry (APR), severity and mortality indicators. We pull the drug severity information and right-join it to our query tables.
```{r}
tbl(con, "drgcodes") %>%
  filter(str_detect(drg_type, "APR")) %>%
  select(subject_id, hadm_id, drg_severity) %>%
  right_join(study_admissions_ccu, by = c("subject_id", "hadm_id")) %>%
  mutate(drg_severity = ifelse(is.na(drg_severity), 1, drg_severity)) %>%
  print() -> study_admissions_ccu
study_admissions_csru <- tbl(con, "drgcodes") %>%
  filter(str_detect(drg_type, "APR")) %>%
  select(subject_id, hadm_id, drg_severity) %>%
  right_join(study_admissions_csru, by = c("subject_id", "hadm_id")) %>%
  mutate(drg_severity = ifelse(is.na(drg_severity), 1, drg_severity)) %>%
  print() -> study_admissions_csru
study_admissions_micu <- tbl(con, "drgcodes") %>%
  filter(str_detect(drg_type, "APR")) %>%
  select(subject_id, hadm_id, drg_severity) %>%
  right_join(study_admissions_micu, by = c("subject_id", "hadm_id")) %>%
  mutate(drg_severity = ifelse(is.na(drg_severity), 1, drg_severity)) %>%
  print() -> study_admissions_micu
study_admissions_sicu <- tbl(con, "drgcodes") %>%
  filter(str_detect(drg_type, "APR")) %>%
  select(subject_id, hadm_id, drg_severity) %>%
  right_join(study_admissions_sicu, by = c("subject_id", "hadm_id")) %>%
  mutate(drg_severity = ifelse(is.na(drg_severity), 1, drg_severity)) %>%
  print() -> study_admissions_sicu
study_admissions_tsicu <- tbl(con, "drgcodes") %>%
  filter(str_detect(drg_type, "APR")) %>%
  select(subject_id, hadm_id, drg_severity) %>%
  right_join(study_admissions_tsicu, by = c("subject_id", "hadm_id")) %>%
  mutate(drg_severity = ifelse(is.na(drg_severity), 1, drg_severity)) %>%
  print() -> study_admissions_tsicu

```

Pull the admission time `admittime`, discharge time `dischtime`, date of birth `dob`, and date of death `dod`. We are interested in the  mortaility rate 30 days after discharge. So we only keep patients who didn't die in hospital.
```{r}
study_admissions_ccu %>%
  left_join(
    select(tbl(con, "admissions"),
           subject_id, hadm_id, admittime, dischtime, hospital_expire_flag
    ), by = c("subject_id", "hadm_id")
  ) %>%
  filter(hospital_expire_flag == 0) %>% # patients who did not die in hospital
  select(-hospital_expire_flag) %>%
  left_join(
    select(tbl(con, "patients"), subject_id, dob, dod),
    by = "subject_id"
  ) %>%
  print(width = Inf) -> study_admissions_ccu

study_admissions_csru %>%
  left_join(
    select(tbl(con, "admissions"),
           subject_id, hadm_id, admittime, dischtime, hospital_expire_flag
    ), by = c("subject_id", "hadm_id")
  ) %>%
  filter(hospital_expire_flag == 0) %>% # patients who did not die in hospital
  select(-hospital_expire_flag) %>%
  left_join(
    select(tbl(con, "patients"), subject_id, dob, dod),
    by = "subject_id"
  ) %>%
  print(width = Inf) -> study_admissions_csru

study_admissions_micu %>%
  left_join(
    select(tbl(con, "admissions"),
           subject_id, hadm_id, admittime, dischtime, hospital_expire_flag
    ), by = c("subject_id", "hadm_id")
  ) %>%
  filter(hospital_expire_flag == 0) %>% # patients who did not die in hospital
  select(-hospital_expire_flag) %>%
  left_join(
    select(tbl(con, "patients"), subject_id, dob, dod),
    by = "subject_id"
  ) %>%
  print(width = Inf) -> study_admissions_micu

study_admissions_sicu %>%
  left_join(
    select(tbl(con, "admissions"),
           subject_id, hadm_id, admittime, dischtime, hospital_expire_flag
    ), by = c("subject_id", "hadm_id")
  ) %>%
  filter(hospital_expire_flag == 0) %>% # patients who did not die in hospital
  select(-hospital_expire_flag) %>%
  left_join(
    select(tbl(con, "patients"), subject_id, dob, dod),
    by = "subject_id"
  ) %>%
  print(width = Inf) -> study_admissions_sicu

study_admissions_tsicu %>%
  left_join(
    select(tbl(con, "admissions"),
           subject_id, hadm_id, admittime, dischtime, hospital_expire_flag
    ), by = c("subject_id", "hadm_id")
  ) %>%
  filter(hospital_expire_flag == 0) %>% # patients who did not die in hospital
  select(-hospital_expire_flag) %>%
  left_join(
    select(tbl(con, "patients"), subject_id, dob, dod),
    by = "subject_id"
  ) %>%
  print(width = Inf) -> study_admissions_tsicu
```

To add `age` (at admission) variable into the table. [The documentation for the patients table](https://mimic.physionet.org/mimictables/patients/) explains that patients of 90 years and older had their ages artificially inflated, so we remove these patients from the analysis.
```{r}
study_admissions_ccu %>%
  mutate(tt_death = date_part("day", dod) - date_part("day", dischtime)) %>%
  mutate(mortality = tt_death <= 30) %>%
  mutate(age = date_part("year", admittime) - date_part("year", dob)) %>%
  filter(age < 90) %>%
  mutate(age = age - ifelse(
    date_part("month", admittime) < date_part("month", dob) |
      (
        date_part("month", admittime) == date_part("month", dob) &
          date_part("day", admittime) < date_part("day", dob)
      ),
    1,
    0
  )) %>%
  select(-admittime, -dischtime, -dob, -dod, -tt_death) %>%
  select(subject_id, hadm_id, age, mortality, everything()) %>%
  print() -> study_admissions_ccu

study_admissions_csru %>%
  mutate(tt_death = date_part("day", dod) - date_part("day", dischtime)) %>%
  mutate(mortality = tt_death <= 30) %>%
  mutate(age = date_part("year", admittime) - date_part("year", dob)) %>%
  filter(age < 90) %>%
  mutate(age = age - ifelse(
    date_part("month", admittime) < date_part("month", dob) |
      (
        date_part("month", admittime) == date_part("month", dob) &
          date_part("day", admittime) < date_part("day", dob)
      ),
    1,
    0
  )) %>%
  select(-admittime, -dischtime, -dob, -dod, -tt_death) %>%
  select(subject_id, hadm_id, age, mortality, everything()) %>%
  print() -> study_admissions_csru

study_admissions_micu %>%
  mutate(tt_death = date_part("day", dod) - date_part("day", dischtime)) %>%
  mutate(mortality = tt_death <= 30) %>%
  mutate(age = date_part("year", admittime) - date_part("year", dob)) %>%
  filter(age < 90) %>%
  mutate(age = age - ifelse(
    date_part("month", admittime) < date_part("month", dob) |
      (
        date_part("month", admittime) == date_part("month", dob) &
          date_part("day", admittime) < date_part("day", dob)
      ),
    1,
    0
  )) %>%
  select(-admittime, -dischtime, -dob, -dod, -tt_death) %>%
  select(subject_id, hadm_id, age, mortality, everything()) %>%
  print() -> study_admissions_micu

study_admissions_sicu %>%
  mutate(tt_death = date_part("day", dod) - date_part("day", dischtime)) %>%
  mutate(mortality = tt_death <= 30) %>%
  mutate(age = date_part("year", admittime) - date_part("year", dob)) %>%
  filter(age < 90) %>%
  mutate(age = age - ifelse(
    date_part("month", admittime) < date_part("month", dob) |
      (
        date_part("month", admittime) == date_part("month", dob) &
          date_part("day", admittime) < date_part("day", dob)
      ),
    1,
    0
  )) %>%
  select(-admittime, -dischtime, -dob, -dod, -tt_death) %>%
  select(subject_id, hadm_id, age, mortality, everything()) %>%
  print() -> study_admissions_sicu

study_admissions_tsicu %>%
  mutate(tt_death = date_part("day", dod) - date_part("day", dischtime)) %>%
  mutate(mortality = tt_death <= 30) %>%
  mutate(age = date_part("year", admittime) - date_part("year", dob)) %>%
  filter(age < 90) %>%
  mutate(age = age - ifelse(
    date_part("month", admittime) < date_part("month", dob) |
      (
        date_part("month", admittime) == date_part("month", dob) &
          date_part("day", admittime) < date_part("day", dob)
      ),
    1,
    0
  )) %>%
  select(-admittime, -dischtime, -dob, -dod, -tt_death) %>%
  select(subject_id, hadm_id, age, mortality, everything()) %>%
  print() -> study_admissions_tsicu

```

Many mortality indicators are missing, due to neither the hospital database nor the social security database having a record of these patients’ deaths. It may be helpful to retain in the analytic table this information on whether deaths were recorded at all, e.g. for validation or sensitivity testing.

Finally, let's merge some demographic information (ethnicity, gender) into our study.
```{r}
tbl(con, "admissions") %>%
  select(subject_id, ethnicity) %>%
  distinct() %>%
  print() -> study_subjects
```
```{r}
tbl(con, "patients") %>%
  select(subject_id, gender) %>%
  distinct() %>%
  full_join(study_subjects, by = "subject_id") %>%
  print() -> study_subjects
```


Let's resolves ome diversity and inconsistency in the `ethnicity` field:
```{r}
unknown_ethnicity <- c(
  "OTHER",
  "UNABLE TO OBTAIN",
  "UNKNOWN/NOT SPECIFIED",
  "MULTI RACE ETHNICITY",
  "PATIENT DECLINED TO ANSWER",
  "UNKNOWN"
)

study_subjects %>%
  collect() %>%
  mutate(ethnic_group = case_when(
    str_detect(ethnicity, "^ASIAN") ~ "ASIAN",
    str_detect(ethnicity, "^BLACK") ~ "BLACK",
    str_detect(ethnicity, "^HISPANIC") ~ "HISPANIC",
    str_detect(ethnicity, "^WHITE") ~ "WHITE",
    ethnicity %in% unknown_ethnicity ~ NA_character_,
    TRUE ~ NA_character_
  )) %>%
  select(subject_id, gender, ethnic_group) %>%
  print() -> study_subjects
```

Some patients are coded as belonging to more than one ethnic group. To resolve these inconsistencies, we define a helper function to pick the modal value from a vector of values in R, which can be used by the `summarize()` function to choose one ethnic group for each patient.
```{r}
most <- function(x) {
  if (all(is.na(x))) return(NA_character_)
  y <- table(x, useNA = "no")
  if (length(which(y == max(y))) > 1) return(NA_character_)
  return(names(y)[which.max(y)])
}

study_subjects %>%
  group_by(subject_id) %>%
  summarize(ethnic_group = most(ethnic_group)) %>%
  ungroup() %>%
  mutate(ethnic_group = ifelse(is.na(ethnic_group), "UNKNOWN", ethnic_group)) %>%
  print() -> subject_ethnic_groups
```
```{r}
study_subjects %>%
  select(subject_id, gender) %>%
  left_join(subject_ethnic_groups, by = "subject_id") %>%
  print() -> study_subjects
```

Now we add the demographic information `gender` and `ethnicity` into our tables:
```{r}
study_admissions_ccu %>%
  left_join(study_subjects, by = "subject_id", copy = TRUE) %>%
  print() -> study_admissions_ccu
study_admissions_csru %>%
  left_join(study_subjects, by = "subject_id", copy = TRUE) %>%
  print() -> study_admissions_csru
study_admissions_micu %>%
  left_join(study_subjects, by = "subject_id", copy = TRUE) %>%
  print() -> study_admissions_micu
study_admissions_sicu %>%
  left_join(study_subjects, by = "subject_id", copy = TRUE) %>%
  print() -> study_admissions_sicu
study_admissions_tsicu %>%
  left_join(study_subjects, by = "subject_id", copy = TRUE) %>%
  print() -> study_admissions_tsicu
```

Close the connection:
```{r}
dbDisconnect(con)
```
# Data Visualization

# Analytics

# Conclusions
