---
title: "30-Day Mortality Condition of Myocardia Infarction Patients Admitted to All Seven Care Units"
author: Yu Chen 305266880
output:
  # ioslides_presentation: default
  html_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This report analyzes the electronic health record (EHR) data MIMIC-III. By working with PostgreSQL database, data visualization, and commonly used analytical methods such as logistic regression and neural network, I try to analyze 30-day mortality condition of myocardia infarction patients admitted to all seven care units. It shows a better picture of hospital conditions and can have a positive impact on decision making.

# Data Preparation

Load database libraries and the tidyverse frontend:
```{r}
library(DBI)
library(RPostgreSQL)
library(tidyverse)
library(lubridate)
```

Credentials for using PostgreSQL database. We are going to use username `postgres` with password `postgres` to access the `mimic` database in the schemee `mimiciii`. 
```{r}
# Load configuration settings
dbdriver <- 'PostgreSQL'
#host  <- '127.0.0.1'
#port  <- '5432'
user  <- 'postgres'
password <- 'postgres'
dbname <- 'mimic'
schema <- 'mimiciii'
# Connect to the database using the configuration settings
con <- dbConnect(RPostgreSQL::PostgreSQL(), 
                 dbname = dbname, 
                 #host = host, 
                 #port = port, 
                 user = user, 
                 password = password)
# Set the default schema
dbExecute(con, paste("SET search_path TO ", schema, sep=" "))
con
```

In this section, we demo how to create a cohort of patients admitted into CCU, CSRU, MICU, NICU, NWARD, SICU and TSICU and were diagnosed with Myocardia Infarction.

First we create a (query) table of patients who were directly admitted into CCU, CSRU, MICU, NICU, NWARD, SICU and TSICU.
```{r}
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "CCU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> ccu_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "CSRU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> csru_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "MICU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> micu_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "NICU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> nicu_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "NWARD") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> nward_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "SICU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> sicu_admissions
tbl(con, "transfers") %>%
  select(subject_id, hadm_id, prev_careunit, curr_careunit) %>%
  filter(is.na(prev_careunit) & curr_careunit == "TSICU") %>%
  select(subject_id, hadm_id) %>%
  distinct() %>%
  print() -> tsicu_admissions
```

Now we want to restrict to Myocardia Infarction Patients. To find all possible ICD-9 codes related to heart attack, we search for string `myocardial infarction` in the `long_title` of table `d_icd_diagnoses`:
```{r}
mi_codes <- tbl(con, "d_icd_diagnoses") %>%
  filter(str_detect(tolower(long_title), "myocardial infarction")) 
```

`diagnoses_icd` table stores the diagnosis of each admission. We use `semi_join()` to keep the rows in `diagnoses_icd` that match the ICD-9 codes related to fever:
```{r}
tbl(con, "diagnoses_icd") %>%
  semi_join(mi_codes, by = "icd9_code") %>%
  print() -> mi_admissions
```

MI may not be listed as the principal diagnosis; as explained in [the documentation for the `patients` table](https://mimic.physionet.org/mimictables/diagnoses_icd/), the `seq_num` field is a priority ranking for the diagnoses generated at the end of stay. In order to focus on patients for whom MI was central to their hospitalization, we will include records with MI in any of the first five diagnosis positions, according to the `seq_num` field. To avoid duplicate admissions, we use `group_by()` and `top_n()` to limit the query to the first MI diagnosis for each admission.
```{r}
mi_admissions <- mi_admissions %>%
  filter(seq_num <= 5) %>%
  group_by(subject_id, hadm_id) %>%
  filter(min_rank(seq_num) <= 1) %>%
  ungroup() %>%
  select(subject_id, hadm_id, icd9_code, seq_num)
```

Now we `inner_join` the table of admissions to all the unit cares and the table of admissions that include MI diagnosis.
```{r}
study_admissions_ccu <- ccu_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_csru <- csru_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_micu <- micu_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_nicu <- nicu_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_nward <- nward_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_sicu <- sicu_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
study_admissions_tsicu <- tsicu_admissions %>%
  inner_join(mi_admissions, by = c("subject_id", "hadm_id")) %>%
  print()
```

We notice that there is no data in table `study_admissions_nicu` and `study_admissions_nward`, which means there are no fever patients addmitted to NICU and NWARD. So we won't consider these two conditions in the following analysis.

Now we create a logical variable indicating the MI is the principal diagonosis or not (according to `seq_num`).
```{r}
study_admissions_ccu <- study_admissions_ccu %>%
  mutate(principal_dx = seq_num == 1) %>%
  select(-seq_num) %>%
  print()
study_admissions_csru <- study_admissions_csru %>%
  mutate(principal_dx = seq_num == 1) %>%
  select(-seq_num) %>%
  print()
study_admissions_micu <- study_admissions_micu %>%
  mutate(principal_dx = seq_num == 1) %>%
  select(-seq_num) %>%
  print()
study_admissions_sicu <- study_admissions_sicu %>%
  mutate(principal_dx = seq_num == 1) %>%
  select(-seq_num) %>%
  print()
study_admissions_tsicu <- study_admissions_tsicu %>%
  mutate(principal_dx = seq_num == 1) %>%
  select(-seq_num) %>%
  print()
```

We want to add information about the severity of patients’ ailments. The `drgcodes` table contains, for `DRG` codes from the All Payers Registry (APR), severity and mortality indicators. We pull the drug severity information and right-join it to our query tables.
```{r}
tbl(con, "drgcodes") %>%
  filter(str_detect(drg_type, "APR")) %>%
  select(subject_id, hadm_id, drg_severity) %>%
  right_join(study_admissions_ccu, by = c("subject_id", "hadm_id")) %>%
  mutate(drg_severity = ifelse(is.na(drg_severity), 1, drg_severity)) %>%
  print() -> study_admissions_ccu
study_admissions_csru <- tbl(con, "drgcodes") %>%
  filter(str_detect(drg_type, "APR")) %>%
  select(subject_id, hadm_id, drg_severity) %>%
  right_join(study_admissions_csru, by = c("subject_id", "hadm_id")) %>%
  mutate(drg_severity = ifelse(is.na(drg_severity), 1, drg_severity)) %>%
  print() -> study_admissions_csru
study_admissions_micu <- tbl(con, "drgcodes") %>%
  filter(str_detect(drg_type, "APR")) %>%
  select(subject_id, hadm_id, drg_severity) %>%
  right_join(study_admissions_micu, by = c("subject_id", "hadm_id")) %>%
  mutate(drg_severity = ifelse(is.na(drg_severity), 1, drg_severity)) %>%
  print() -> study_admissions_micu
study_admissions_sicu <- tbl(con, "drgcodes") %>%
  filter(str_detect(drg_type, "APR")) %>%
  select(subject_id, hadm_id, drg_severity) %>%
  right_join(study_admissions_sicu, by = c("subject_id", "hadm_id")) %>%
  mutate(drg_severity = ifelse(is.na(drg_severity), 1, drg_severity)) %>%
  print() -> study_admissions_sicu
study_admissions_tsicu <- tbl(con, "drgcodes") %>%
  filter(str_detect(drg_type, "APR")) %>%
  select(subject_id, hadm_id, drg_severity) %>%
  right_join(study_admissions_tsicu, by = c("subject_id", "hadm_id")) %>%
  mutate(drg_severity = ifelse(is.na(drg_severity), 1, drg_severity)) %>%
  print() -> study_admissions_tsicu

```

Pull the admission time `admittime`, discharge time `dischtime`, date of birth `dob`, and date of death `dod`. We are interested in the mortaility rate 30 days after discharge. So we only keep patients who didn't die in hospital.
```{r}
study_admissions_ccu %>%
  left_join(
    select(tbl(con, "admissions"),
           subject_id, hadm_id, admittime, dischtime, hospital_expire_flag
    ), by = c("subject_id", "hadm_id")
  ) %>%
  filter(hospital_expire_flag == 0) %>% # patients who did not die in hospital
  select(-hospital_expire_flag) %>%
  left_join(
    select(tbl(con, "patients"), subject_id, dob, dod),
    by = "subject_id"
  ) %>%
  print(width = Inf) -> study_admissions_ccu

study_admissions_csru %>%
  left_join(
    select(tbl(con, "admissions"),
           subject_id, hadm_id, admittime, dischtime, hospital_expire_flag
    ), by = c("subject_id", "hadm_id")
  ) %>%
  filter(hospital_expire_flag == 0) %>% # patients who did not die in hospital
  select(-hospital_expire_flag) %>%
  left_join(
    select(tbl(con, "patients"), subject_id, dob, dod),
    by = "subject_id"
  ) %>%
  print(width = Inf) -> study_admissions_csru

study_admissions_micu %>%
  left_join(
    select(tbl(con, "admissions"),
           subject_id, hadm_id, admittime, dischtime, hospital_expire_flag
    ), by = c("subject_id", "hadm_id")
  ) %>%
  filter(hospital_expire_flag == 0) %>% # patients who did not die in hospital
  select(-hospital_expire_flag) %>%
  left_join(
    select(tbl(con, "patients"), subject_id, dob, dod),
    by = "subject_id"
  ) %>%
  print(width = Inf) -> study_admissions_micu

study_admissions_sicu %>%
  left_join(
    select(tbl(con, "admissions"),
           subject_id, hadm_id, admittime, dischtime, hospital_expire_flag
    ), by = c("subject_id", "hadm_id")
  ) %>%
  filter(hospital_expire_flag == 0) %>% # patients who did not die in hospital
  select(-hospital_expire_flag) %>%
  left_join(
    select(tbl(con, "patients"), subject_id, dob, dod),
    by = "subject_id"
  ) %>%
  print(width = Inf) -> study_admissions_sicu

study_admissions_tsicu %>%
  left_join(
    select(tbl(con, "admissions"),
           subject_id, hadm_id, admittime, dischtime, hospital_expire_flag
    ), by = c("subject_id", "hadm_id")
  ) %>%
  filter(hospital_expire_flag == 0) %>% # patients who did not die in hospital
  select(-hospital_expire_flag) %>%
  left_join(
    select(tbl(con, "patients"), subject_id, dob, dod),
    by = "subject_id"
  ) %>%
  print(width = Inf) -> study_admissions_tsicu
```

To add `age` (at admission) variable into the table. [The documentation for the patients table](https://mimic.physionet.org/mimictables/patients/) explains that patients of 90 years and older had their ages artificially inflated, so we remove these patients from the analysis.
```{r}
study_admissions_ccu %>%
  mutate(tt_death = date_part("day", dod) - date_part("day", dischtime)) %>%
  mutate(mortality = tt_death <= 30) %>%
  mutate(age = date_part("year", admittime) - date_part("year", dob)) %>%
  filter(age < 90) %>%
  mutate(age = age - ifelse(
    date_part("month", admittime) < date_part("month", dob) |
      (
        date_part("month", admittime) == date_part("month", dob) &
          date_part("day", admittime) < date_part("day", dob)
      ),
    1,
    0
  )) %>%
  select(-admittime, -dischtime, -dob, -dod, -tt_death) %>%
  select(subject_id, hadm_id, age, mortality, everything()) %>%
  print() -> study_admissions_ccu

study_admissions_csru %>%
  mutate(tt_death = date_part("day", dod) - date_part("day", dischtime)) %>%
  mutate(mortality = tt_death <= 30) %>%
  mutate(age = date_part("year", admittime) - date_part("year", dob)) %>%
  filter(age < 90) %>%
  mutate(age = age - ifelse(
    date_part("month", admittime) < date_part("month", dob) |
      (
        date_part("month", admittime) == date_part("month", dob) &
          date_part("day", admittime) < date_part("day", dob)
      ),
    1,
    0
  )) %>%
  select(-admittime, -dischtime, -dob, -dod, -tt_death) %>%
  select(subject_id, hadm_id, age, mortality, everything()) %>%
  print() -> study_admissions_csru

study_admissions_micu %>%
  mutate(tt_death = date_part("day", dod) - date_part("day", dischtime)) %>%
  mutate(mortality = tt_death <= 30) %>%
  mutate(age = date_part("year", admittime) - date_part("year", dob)) %>%
  filter(age < 90) %>%
  mutate(age = age - ifelse(
    date_part("month", admittime) < date_part("month", dob) |
      (
        date_part("month", admittime) == date_part("month", dob) &
          date_part("day", admittime) < date_part("day", dob)
      ),
    1,
    0
  )) %>%
  select(-admittime, -dischtime, -dob, -dod, -tt_death) %>%
  select(subject_id, hadm_id, age, mortality, everything()) %>%
  print() -> study_admissions_micu

study_admissions_sicu %>%
  mutate(tt_death = date_part("day", dod) - date_part("day", dischtime)) %>%
  mutate(mortality = tt_death <= 30) %>%
  mutate(age = date_part("year", admittime) - date_part("year", dob)) %>%
  filter(age < 90) %>%
  mutate(age = age - ifelse(
    date_part("month", admittime) < date_part("month", dob) |
      (
        date_part("month", admittime) == date_part("month", dob) &
          date_part("day", admittime) < date_part("day", dob)
      ),
    1,
    0
  )) %>%
  select(-admittime, -dischtime, -dob, -dod, -tt_death) %>%
  select(subject_id, hadm_id, age, mortality, everything()) %>%
  print() -> study_admissions_sicu

study_admissions_tsicu %>%
  mutate(tt_death = date_part("day", dod) - date_part("day", dischtime)) %>%
  mutate(mortality = tt_death <= 30) %>%
  mutate(age = date_part("year", admittime) - date_part("year", dob)) %>%
  filter(age < 90) %>%
  mutate(age = age - ifelse(
    date_part("month", admittime) < date_part("month", dob) |
      (
        date_part("month", admittime) == date_part("month", dob) &
          date_part("day", admittime) < date_part("day", dob)
      ),
    1,
    0
  )) %>%
  select(-admittime, -dischtime, -dob, -dod, -tt_death) %>%
  select(subject_id, hadm_id, age, mortality, everything()) %>%
  print() -> study_admissions_tsicu

```

Many mortality indicators are missing, due to neither the hospital database nor the social security database having a record of these patients’ deaths. It may be helpful to retain in the analytic table this information on whether deaths were recorded at all, e.g. for validation or sensitivity testing.

Finally, let's merge some demographic information (ethnicity, gender) into our study.
```{r}
tbl(con, "admissions") %>%
  select(subject_id, ethnicity) %>%
  distinct() %>%
  print() -> study_subjects
```
```{r}
tbl(con, "patients") %>%
  select(subject_id, gender) %>%
  distinct() %>%
  full_join(study_subjects, by = "subject_id") %>%
  print() -> study_subjects
```


Let's resolves ome diversity and inconsistency in the `ethnicity` field:
```{r}
unknown_ethnicity <- c(
  "OTHER",
  "UNABLE TO OBTAIN",
  "UNKNOWN/NOT SPECIFIED",
  "MULTI RACE ETHNICITY",
  "PATIENT DECLINED TO ANSWER",
  "UNKNOWN"
)

study_subjects %>%
  collect() %>%
  mutate(ethnic_group = case_when(
    str_detect(ethnicity, "^ASIAN") ~ "ASIAN",
    str_detect(ethnicity, "^BLACK") ~ "BLACK",
    str_detect(ethnicity, "^HISPANIC") ~ "HISPANIC",
    str_detect(ethnicity, "^WHITE") ~ "WHITE",
    ethnicity %in% unknown_ethnicity ~ NA_character_,
    TRUE ~ NA_character_
  )) %>%
  select(subject_id, gender, ethnic_group) %>%
  print() -> study_subjects
```

Some patients are coded as belonging to more than one ethnic group. To resolve these inconsistencies, we define a helper function to pick the modal value from a vector of values in R, which can be used by the `summarize()` function to choose one ethnic group for each patient.
```{r}
most <- function(x) {
  if (all(is.na(x))) return(NA_character_)
  y <- table(x, useNA = "no")
  if (length(which(y == max(y))) > 1) return(NA_character_)
  return(names(y)[which.max(y)])
}

study_subjects %>%
  group_by(subject_id) %>%
  summarize(ethnic_group = most(ethnic_group)) %>%
  ungroup() %>%
  mutate(ethnic_group = ifelse(is.na(ethnic_group), "UNKNOWN", ethnic_group)) %>%
  print() -> subject_ethnic_groups
```
```{r}
study_subjects %>%
  select(subject_id, gender) %>%
  left_join(subject_ethnic_groups, by = "subject_id") %>%
  print() -> study_subjects
```

Now we add the demographic information `gender` and `ethnicity` into our tables:
```{r}
study_admissions_ccu %>%
  left_join(study_subjects, by = "subject_id", copy = TRUE) %>%
  print() -> study_admissions_ccu
study_admissions_csru %>%
  left_join(study_subjects, by = "subject_id", copy = TRUE) %>%
  print() -> study_admissions_csru
study_admissions_micu %>%
  left_join(study_subjects, by = "subject_id", copy = TRUE) %>%
  print() -> study_admissions_micu
study_admissions_sicu %>%
  left_join(study_subjects, by = "subject_id", copy = TRUE) %>%
  print() -> study_admissions_sicu
study_admissions_tsicu %>%
  left_join(study_subjects, by = "subject_id", copy = TRUE) %>%
  print() -> study_admissions_tsicu
```

I load the entire tables into the memory we may use:
```{r}
ccu <- collect(study_admissions_ccu)
csru <- collect(study_admissions_csru)
micu <- collect(study_admissions_micu)
sicu <- collect(study_admissions_sicu)
tsicu <- collect(study_admissions_tsicu)
```

# Data Visualization

```{r plot}
library(shape)
library(diagram)

# set margins and multiplot
par(mfrow = c(1, 1))
par(mar = c(0, 0, 0, 0))

# initialise a plot device
openplotmat()

# position of boxes
# 1st column indicates x axis position between 0 and 1
# 2nd column indicates y axis position between 0 and 1
# automatically assigns vertical position
num_of_boxes <- 6
auto_coords = coordinates(num_of_boxes)
vert_pos = rev(auto_coords[,1])
box_pos <- matrix(nrow = num_of_boxes, ncol = 2, data = 0)
box_pos[1,] = c(0.30, vert_pos[1]) # 1st box
box_pos[2,] = c(0.30, vert_pos[2]) # 2nd box
box_pos[3,] = c(0.70, vert_pos[3]) # 3rd box
box_pos[4,] = c(0.30, vert_pos[4]) # etc...
box_pos[5,] = c(0.30, vert_pos[5])
box_pos[6,] = c(0.30, vert_pos[6])

# content of boxes
box_content <- matrix(nrow = num_of_boxes, ncol = 1, data = 0)
box_content[1] = "Patients admitted to all care units" # 1st box
box_content[2] = "Patients diagnosed with Myocardia Infarction" # 2nd box
box_content[3] = "Add round inf about patients" # 3rd box
box_content[4] = "Calculate death rate and population" # etc...
box_content[5] = "Detailed analysis of patients admitted to CCU and SICU"
box_content[6] = "Add age, severity, gender and ethnicity"


# adjust the size of boxes to fit content
box_x <- c(0.25, 0.25, 0.25, 0.25, 0.30, 0.25, 0.25)
box_y <- c(0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05)

# Draw the arrows
straightarrow(from = box_pos[1,], to = box_pos[2,], lwd = 1)  
straightarrow(from = c(box_pos[2,1],box_pos[3,2]), to = box_pos[3,], lwd = 1)  
straightarrow(from = box_pos[2,], to = box_pos[4,], lwd = 1)  
straightarrow(from = box_pos[4,], to = box_pos[5,], lwd = 1)  
straightarrow(from = box_pos[5,], to = box_pos[6,], lwd = 1) 

# Draw the boxes
for (i in 1:num_of_boxes) {
  textrect(mid = box_pos[i,], radx = box_x[i], rady = box_y[i], 
           lab = box_content[i], 
           shadow.col = "grey")
  }
```

In this part, we calculate the death rate in all unit cares to see the difference.
```{r}
tbl <- as_tibble(data.frame(care_unit = c("ccu", "csru", "micu", "sicu", "tsicu"))) %>%
  mutate(deathrate = c(sum(!is.na(ccu$mortality))/nrow(ccu), sum(!is.na(csru$mortality))/nrow(csru), sum(!is.na(micu$mortality))/nrow(micu), sum(!is.na(sicu$mortality))/nrow(sicu), sum(!is.na(tsicu$mortality))/nrow(tsicu))) %>%
  mutate(population = c(nrow(ccu), nrow(csru), nrow(micu), nrow(sicu), nrow(tsicu))) %>%
  mutate(number_of_death = c(sum(!is.na(ccu$mortality)), sum(!is.na(csru$mortality)), sum(!is.na(micu$mortality)), sum(!is.na(sicu$mortality)), sum(!is.na(tsicu$mortality)))) %>%
  print()
```

We can see the highest death rate happens in SICU, and highest number of patients were admitted to CCU, so we analyze the detailed information of passed patients transfered to SICU and CCU in the following parts.

## CCU

- Age
```{r}
ccu %>%
  ggplot(aes(x = age)) +
  geom_freqpoly(bins = 20) +
  labs(x = "Age (CCU)")
```

As shown in the figure above, the majority of myocardia infarction patients admitted to SICU are between 55 to 70 years old, which makes sence because elderly people may have higher risk in getting heart attack.

Then we consider people who died within 30 days after discharge.
```{r}
ccu %>%
  filter(mortality == TRUE) %>%
  ggplot(aes(x = age)) +
  geom_freqpoly(bins = 20) +
  labs(x = "Age (CCU)")
```

A better picture is shown below.
```{r}
ccu %>%
  filter(mortality == TRUE) %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = factor(1), y = age)) +
  labs(x = "", y = "Age") + 
  scale_y_log10()
```

It's clearly that the majority of myocardia infarction patients admitted to SICU and died within 30 days is about 72 years old.

- Severity

In this section, we want to discuss the severity of passed patients’ ailments when they leave the hospital.
```{r}
ccu %>%
  filter(mortality == TRUE) %>%
  ggplot() +
  geom_bar(mapping = aes(x = drg_severity)) +
  labs(x = "Severity")
```

We can see that most of the passed patients leave the hospital when their diagnosis are not very severe and less than half of the patients are highly severe.

- Gender

```{r}
ccu %>%
  filter(mortality == TRUE) %>%
  ggplot() +
  geom_bar(mapping = aes(x = gender)) +
  labs(x = "Gender")
```

In this case, we know that more male patients passed within 30 days, it's because the number of males patients is larger than that of female patients in the first place.

-Ethnicity
```{r}
ccu %>%
  filter(mortality == TRUE) %>%
  filter(ethnic_group != "UNKNOWN") %>%
  ggplot() +
  geom_bar(mapping = aes(x = ethnic_group)) +
  labs(x = "Ethnicity")
```

Ignoring the unknown ethnicity, we can see that the majority of passed patients are white, only a few patients are Asian or Hispanic.

## SICU

- Age
```{r}
sicu %>%
  ggplot(aes(x = age)) +
  geom_freqpoly(bins = 20) +
  labs(x = "Age (SICU)")
```

As shown in the figure above, the majority of myocardia infarction patients admitted to SICU are between 60 to 80 years old, which makes sence because elderly people may have higher risk in getting heart attack.

Then we consider people who died within 30 days after discharge.
```{r}
sicu %>%
  filter(mortality == TRUE) %>%
  ggplot(aes(x = age)) +
  geom_freqpoly(bins = 20) +
  labs(x = "Age")
```

A better picture is shown below.
```{r}
sicu %>%
  filter(mortality == TRUE) %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = factor(1), y = age)) +
  labs(x = "", y = "Age") + 
  scale_y_log10()
```

It's clearly that the majority of myocardia infarction patients admitted to SICU and died within 30 days is about 72 years old.

- Severity

In this section, we want to discuss the severity of passed patients’ ailments when they leave the hospital.
```{r}
sicu %>%
  filter(mortality == TRUE) %>%
  ggplot() +
  geom_bar(mapping = aes(x = drg_severity)) +
  labs(x = "Severity")
```

We can see that most of the passed patients leave the hospital when their diagnosis are not very severe and less than a quarter of the patients are highly severe.

- Gender

```{r}
sicu %>%
  filter(mortality == TRUE) %>%
  ggplot() +
  geom_bar(mapping = aes(x = gender)) +
  labs(x = "Gender")
```

In this case, we know that more male patients passed within 30 days, it's because the number of males patients is larger than that of female patients in the first place.

-Ethnicity
```{r}
sicu %>%
  filter(mortality == TRUE) %>%
  filter(ethnic_group != "UNKNOWN") %>%
  ggplot() +
  geom_bar(mapping = aes(x = ethnic_group)) +
  labs(x = "Ethnicity")
```

Ignoring the unknown ethnicity, we can see that the majority of passed patients are white.
```{r}
dbDisconnect(con)
```


# Analytics

```{r}
library(keras)
install_keras()
x_train = x_matrix[1:1000,]
y_train = y_matrix[1:1000,]
x_test = x_matrix[1000:1209,]
y_test = y_matrix[1000:1209,]
dim(x_train)
dim(y_train)
```

```{r}
model <- keras_model_sequential() 
model %>% 
  layer_dense(units = 50, activation = 'relu', input_shape = c(5))
summary(model)
```

Compile the model with appropriate loss function, optimizer, and metrics:
```{r}
model %>% 
  compile(
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)
```

### Training and validation

```{r}
system.time({
history <- model %>% fit(
  x_train, y_train, 
  epochs = 30, batch_size = 128, 
  validation_split = 0.2
)
})
plot(history)
```

### Testing

Evaluate model performance on the test data:
```{r}
model %>% evaluate(x_test, y_test)
```
Generate predictions on new data:
```{r}
model %>% predict_classes(x_test) %>% head()
```

# Conclusions

With the analysis above, we learn that CCU has the largest population of myocardia infarction patients (1675) and 429 of them passed within 30 days after discharge, where the deathrate is 25.61%. The highest deathrate presents itself in case of SICU, i.e., 46.08%. CSRU, MICU, and TSICU have deathrate of 17.16%, 41.13% and 27.45% respectively, while there is no data of  myocardia infarction patients admitted to NICU and NWARD. In this assignment, I detailedly analyse the data of CCU and SICU. I find that most of passed patients are about 72 years old. And male patients are more than female patients, but it may because the initial male patients are more than female patients. Also, I find that most of the passed patients are slightly severe when they leave the hospital, which should draw family members' attention to care of patients after discharge. In the data we study, the marjority of patients are white, which may because the data are collected mainly in white community.

