---
title: "Biostat 203B Homework 2"
subtitle: Due Feb 7 @ 11:59PM
output: html_document
author: YU CHEN (305266880)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

Use tidyverse (ggpot2, dplyr) to explore the [MIMIC-III](https://mimic.physionet.org) data introduced in [homework 1](https://ucla-biostat203b-2020winter.github.io/hw/hw1/hw1.html).
```{r}
library(tidyverse)
library(lubridate)
```

## Q1

Demographic information of the patients admitted into hospital is available in `ADMISSION.csv`. See <https://mimic.physionet.org/mimictables/admissions/> for details of each field in this file. Summarize following variables using appropriate graphs:   

  - admission year  
- admission month  
- admission week day  
- admission hour  
- length of hospital stay  
- admission type  
- number of admissions per patient  
- admission location  
- insurance  
- language  
- religion  
- martial status  
- ethnicity  
- death 

Note it is possible that one patient (uniquely identified by the `SUBJECT_ID`) is admitted into hospital multiple times. When summarizing some demographic information, it makes sense to summarize based on only unique patients. 

**SOLUTION**

Import data first.
```{r}
admission <- read_csv("/home/203bdata/mimic-iii/ADMISSIONS.csv",
                      col_types = cols(ROW_ID = col_integer(),
                                       SUBJECT_ID = col_integer(),
                                       HADM_ID = col_integer(),
                                       ADMITTIME = col_datetime(),
                                       DISCHTIME = col_datetime(),
                                       HOSPITAL_EXPIRE_FLAG = col_logical(),
                                       HAS_CHARTEVENTS_DATA = col_logical()))
admission
```

- Admission Year
```{r}
admission %>%
  ggplot(aes(x = ADMITTIME)) + 
  geom_freqpoly(binwidth = 31536000) + 
  labs(x = "Admission year")
```

- Admission Month
```{r}
admission %>%
  mutate(adm_month = month(ADMITTIME, label = TRUE)) %>%
  ggplot() + 
  geom_bar(mapping = aes(x = adm_month)) + 
  labs(x = 'Admission month')
```

- Admission week day
```{r}
admission %>%
  mutate(adm_wday = wday(ADMITTIME, label = TRUE)) %>%
  ggplot() + 
  geom_bar(mapping = aes(x = adm_wday)) + 
  labs(x = 'Admission week day')
```
We can see from the graph that weekdays are busier than weekends.

- Admission hour
```{r}
admission %>%
  mutate(adm_hour = hour(ADMITTIME)) %>%
  ggplot() + 
  geom_bar(mapping = aes(x = adm_hour)) + 
  labs(x = 'Admission hour')
```
We can see that the 7 am is the busiest time of a day, probably because most people wait and go to hospital the minute it open. Also, hospitals are quite busy in the afternoon as well.

- Length of hospital stay
```{r}
admission %>%
  mutate(hospital_stay = as.numeric(as.duration(DISCHTIME - ADMITTIME) / 86400)) %>%
  ggplot(aes(x = hospital_stay)) +
  geom_freqpoly(bins = 50) +
  labs(x = "Length of hospital stay (days)")
```
A better picture of the pattern of length of hospital stay is shown below.
```{r}
admission %>%
  mutate(hospital_stay = as.numeric(as.duration(DISCHTIME - ADMITTIME) / 86400)) %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = factor(1), y = hospital_stay)) +
  labs(x = "", y = "Length of hospital stay (days)") + 
  scale_y_log10()
```
It is obvious that most patients stay in hospital for around a week.

- Admission type 
```{r}
admission %>%
  ggplot() + 
  geom_bar(mapping = aes(x = ADMISSION_TYPE)) + 
  labs(x = "Admission type")
```

- Number of admissions per patient
!!!

- Admission location
```{r}
admission %>%
  ggplot() + 
  geom_bar(mapping = aes(x = ADMISSION_LOCATION)) + 
  labs(x = "Admission location")
```


## Q2

Link the data in `ADMISSION.csv` and `PATIENTS.csv` (<https://mimic.physionet.org/mimictables/patients/>) and summarize following variables using appropriate graphs:  

- gender  
- age at admission 

## Q3

`ICUSTAYS.csv` (<https://mimic.physionet.org/mimictables/icustays/>) contains data about Intensive Care Units (ICU) stays. Summarize following variables using appropriate graphs:  

- length of ICU stay  
- first ICU unit  
- gender  
- age  

## Q4 

`CHARTEVENTS.csv` (<https://mimic.physionet.org/mimictables/chartevents/>) contains all the charted data available for a patient. During their ICU stay, the primary repository of a patient’s information is their electronic chart. The `ITEMID` variable indicates a single measurement type in the database. The `VALUE` variable is the value measured for `ITEMID`. 

`D_ITEMS.csv` (<https://mimic.physionet.org/mimictables/d_items/>) is the dictionary for the `ITEMID` in `CHARTEVENTS.csv`. Find potential values of `ITEMID` that correspond to systolic blood pressure, i.e., `LABEL` contains the string `systolic`. 

Compile a tibble that contains the first ICU stay of unique patients, with the patient's demographic information, the first systolic blood pressure measurement during ICU stay, and whether the patient died within 30 days of hospitcal admission.